---
title: "Site selection for Rye Nørskov"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(raster)
# if GeoStratR is not installed run devtools::install_github("Sustainscapes/GeoStratR")
library(GeoStratR)
library(sf)
```

## Datasets to be used

In order to include Vegetation complexity and wetness in the startification the following layers were used:

* vegetation height ("O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/canopy_height")
* vegetation density ("O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/vegetation_density")
* vegetation openness (""O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/openness_mean")
* TWI (""O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/twi")

Which resulted in this stack

```{r}
Vars <- readRDS("Variables.rds")
plot(Vars)
```


## Packages used

The `raster` package was used for layer processing, sf for managing shapefiles and the package `GeoStratR` was used for the stratification of the site,

## Raster preparation

The preparation of rasters was made in the 'Prepare_rasters.r' in order to get all the rasters in the same resolution and crs 

## Stratification

The `Stratify` function from GeoSratR was used in order to test the best stratification from 2 to 10 groups, with the following results

```{r, cache = TRUE, message=FALSE, warning=FALSE}
set.seed(2020)
Results <- GeoStratR::Stratify(Vars)
```

In the graph bellow we can see that the number of classes that best captures the variablity is `r Results$Results %>% dplyr::filter(calinski == max(calinski)) %>% pull(n_groups)` as seen in the following graph

```{r}
ggplot(Results$Results, aes(x = n_groups, y = calinski)) + geom_point() + geom_path() + theme_bw() + labs(x = "Number of clusters", y = "Calinski index")
```


The Resulting raster of classes is the following:

```{r}
RyeNoskov <- read_sf("ShapeFiles/RyeNoerskov.shp") %>% 
  st_transform(crs = "+proj=utm +zone=32 +ellps=GRS80 +units=m +no_defs")


FinalStack <- Results$FinalStack

FinalStackDF <- FinalStack %>% 
  as("SpatialPixelsDataFrame") %>% 
  as.data.frame() %>% 
  mutate(Class = case_when(vegetation_density == 1 ~ "A",
                           vegetation_density == 2 ~ "B",
                           vegetation_density == 3 ~ "C"))

ggplot() + 
  geom_raster(data = FinalStackDF, aes(x = x, y = y, fill = Class)) +
  geom_sf(data = RyeNoskov, alpha = 0) +
  labs(x = NULL, y = NULL) +
  theme_bw()
```

Which leads to the following number of cells per class:

```{r}
FinalStackDF %>% group_by(Class) %>% summarise(n = n()) %>% knitr::kable()
```


## Sampling desing

With the following code we will generate 20 random points for sampling withing each class:

```{r, echo=T}
#set seed for reproducibility
set.seed(2021)
# Take the final raster
Sampling <- FinalStackDF %>% 
  ## Divide it by classes
  group_split(Class) %>% 
  ## get 20 random points
  purrr::map(slice_sample, n = 20) %>% 
  ## join everything
  purrr::reduce(bind_rows) %>% 
  ## Select the relevant variables
  dplyr::select(x,y, Class) %>% 
  ## transform to a spatial object
  st_as_sf(coords = c("x", "y"), crs = "+proj=utm +zone=32 +ellps=GRS80 +units=m +no_defs")
```

Which can be seen here:

```{r}
ggplot() + 
  geom_sf(data = RyeNoskov) +
  geom_sf(data = Sampling, aes(color = Class))
  labs(x = NULL, y = NULL) +
  theme_bw()
```

The sampling points are available in the `Sampling` folder

```{r}
dir.create("Sampling")

write_sf(Sampling, "Sampling/Sampling.shp")
```

