---
title: "Site selection for Rye Nørskov"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(raster)
# if GeoStratR is not installed run devtools::install_github("Sustainscapes/GeoStratR")
library(GeoStratR)
library(sf)
```

## Datasets to be used

In order to include Vegetation complexity and wetness in the startification the following layers were used:

* vegetation height ("O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/canopy_height")
* vegetation density ("O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/vegetation_density")
* vegetation openness (""O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/openness_mean")
* TWI (""O:/Nat_Ecoinformatics-tmp/au634851/dk_lidar_backup_2021-06-28/twi")

Which resulted in this stack

```{r}
Vars <- readRDS("Variables.rds")
plot(Vars)
```


## Packages used

The `raster` package was used for layer processing, sf for managing shapefiles and the package `GeoStratR` was used for the stratification of the site,

## Raster preparation

The preparation of rasters was made in the 'Prepare_rasters.r' in order to get all the rasters in the same resolution and crs 

## Stratification

The `Stratify` function from GeoSratR was used in order to test the best stratification from 2 to 10 groups, with the following results

```{r, cache = TRUE, message=FALSE, warning=FALSE}
set.seed(2020)
Results <- GeoStratR::Stratify(Vars)
```

In the graph bellow we can see that the number of classes that best captures the variablity is `r Results$Results %>% dplyr::filter(calinski == max(calinski)) %>% pull(n_groups)` as seen in the following graph

```{r}
ggplot(Results$Results, aes(x = n_groups, y = calinski)) + geom_point() + geom_path() + theme_bw() + labs(x = "Number of clusters", y = "Calinski index")
```


